Inattentive responding can induce spurious associations between task behavior and symptom measures

### 2022-11-16 manuscripts/main.tex
dan's edits

Here we wish to draw special attention to an underappreciated feature of online psychiatric research using self-report symptom surveys. In such surveys, participants rate their endorsement of various psychiatric symptoms and, since most individuals in the general population tend to endorse no or few symptoms in many symptom domains, the resulting ground-truth symptom distributions tend to be heavily positively skewed \cite{lowe2008validation, tomitaka2018distributional}. In this situation, the assumption that C/IE responding merely increases unsystematic measurement noise becomes untenable. Because of the positive skew in the ground-truth symptom distribution, participants who respond carelessly to the symptom survey are more likely to report higher levels of symptom endorsement relative to participants who complete the survey accurately \cite{chandler2020participant, ophir2020turker, king2018random}. Consequently, unless C/IE survey responses are carefully identified and removed, a considerable proportion of putatively symptomatic individuals in an online sample may, in fact, be participants who have not engaged with the experiment with sufficient attention or effort.

@@

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low. In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7; see Figure \ref{fig:distributions}). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct. Permutation testing confirmed that observed mean-shifts in symptom scores for C/IE participants were statistically significant for the majority of symptom measures (Table \ref{tab:validation}).

@@

Second, we quantified the degree to which participants responded to self-report symptom surveys in a stereotyped fashion; that is, we determined if participants exhibited patterns in their responses that were independent of the contents of the survey items. We fit a random-intercept item factor-analysis model \cite{maydeu2006random} to self-report data (see Methods), and for each participant we estimated an intercept parameter that quantified their bias towards using responses on the left or right side of the response scale, regardless of what that response signifies for a particular self-report measure (e.g., low on one symptom scale versus high on another). We observed a credible difference between the average value of this intercept for the two groups ($\Delta \text{intercept} = 0.67, 95\% \ \text{HDI} = [-0.78, -0.55]$), such that C/IE participants were biased towards using the right-half of survey response options. This translates to a tendency to endorse \emph{more severe} symptoms on the 7-up/7-down and GAD-7 scales (where the rightmost options indicate greater frequency of symptoms) but \emph{less severe} symptoms on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms) despite these inventories measuring strongly correlated constructs (i.e. depression and anhedonia, anxiety and behavioral inhibition).

@@

\begin{figure}[!t]
\includegraphics[width=17cm]{../figures/main_03a.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Absolute Spearman rank correlations between task behavior (y-axis) and symptom measures (x-axis) under different regimes of data screening and participant exclusions. Only statistically significant correlations are shown ($p<0.05$ not corrected for multiple comparisons; all correlations are shown in the Supplement). Black Xs indicate significant correlations ablated under screening. No Screening = no exclusions; Accuracy Only = exclusions based on chance-level performance in the reversal-learning task; Infrequency Only = exclusions based on invalid or improbable responses to infrequency items; Both Types = exclusions based on the previous two measures. Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
\label{fig:correlations}
\end{figure}

@@

Next, we investigated how spurious correlations depended on sample size. To do so, we performed a bootstrapping analysis where we held fixed the proportion of participants engaging in C/IE responding (i.e. 5\%, 10\%, 15\%, 20\%) and increased the total number of participants. Across all analyses, we measured the correlation and between the 7-down depression scale and learning-rate asymmetry ($\kappa$), which we previously identified as likely exhibiting a spurious association.

The outputs of the bootstrapping analysis are presented in Figure \ref{fig:bootstrap}. We found that, although estimated correlation magnitudes were independent of sample size (x-axis, left panel), the absolute magnitude of the behavior-symptom correlation increased with the proportion of C/IE participants (different coloured circles, left panel). Crucially, we found false-positive rates for spurious correlations \emph{increased} with increases in sample size in our data for all but the smallest rates of C/IE responding (right panel). This runs counter to a common assumption that larger sample sizes are protective against spurious correlations because they serve to mitigate measurement error. Although this assumption is correct for unsystematic measurement error, it no longer holds in the regime of systematic measurement error (where larger sample sizes reduce the variance of estimates, but do not alter their bias). Instead, our results suggest that, holding the proportion of C/IE responding constant, the false-positive rate for behavioral-symptom correlations will become increasingly inflated as the sample size increases.

@@

Notwithstanding the above, one response to this legitimate concern is to take a graded approach to screening and excluding participants \cite{Kim2018-ev}. That is, participants could be screened with respect to a multitude of measures and only the consistently flagged participants be removed, thereby reducing the risk of inducing bias. Another possibility is to use sensitivity analysis as an alternative to exclusion, testing whether full-sample observed correlations are robust to the exclusion of participants flagged by measures of C/IE responding. We note that the strict screening approach used in the present study did not preclude us from identifying symptomatic participants or behavior-symptom correlations. Indeed, we found in our sample roughly 10\% of participants endorsing symptoms consistent with clinical levels of depression, and approximately 20\% consistent with clinical levels of acute anxiety. These estimates are within the realm of epidemiological norms \cite{kessler2012twelve, lowe2008validation, yarrington2021impact}. We also observed some positive correlations between anxiety and choice behavior that were consistent with effects found in previous literature \cite{huang2017computational, harle2017anhedonia, garrett2018updating}. For example, we found higher lose-shift rates and higher learning rates following negative prediction errors correlated with self-reported anxiety. This suggests that the screening methods we employed were not so aggressive as to attenuate behavior-symptom correlations that would be expected from the literature.

@@

    \item When collecting self-report questionnaire data, include attention-check items that flag participants who may be engaging in C/IE responding. We recommend following best-practice guidelines in using infrequency-item checks rather than instructed-item checks \cite{barends2019noncompliant, thomas2017validity, hauser2016attentive}. Participants flagged by suspicious responses on attention-check items should either be excluded from further analysis, or assessed using sensitivity analyses to ensure that observed full-sample correlations are robust to their exclusion.

@@

To characterize common data screening practices in online computational psychiatry studies, we performed a narrative literature review \cite{grant2009typology}. We identified studies for inclusion through searches on Google Scholar using permutations of query terms related to online labour platforms (e.g. ``mechanical turk", ``prolific", ``online"), experimental paradigms (e.g. ``experiment", ``cognitive control", ``reinforcement learning"), and symptom measures (e.g. ``psychiatry", ``mental illness", ``depression"). We included studies that (a) recruited participants online through a labour platform, (b) measured behavior on at least one experimental task, and (c) measured responses on at least one self-report symptom measure. Through this approach, we identified for inclusion 49 studies spanning 2015 through 2020.

Two of the authors (S.Z., D.B.) then evaluated whether and how each of these studies performed data quality screening for both the collected task and self-report data. Specifically, we confirmed whether a study had performed a particular type of data screening, where the categories were themselves chosen based on commonalities in methods which emerged across the studies. In addition, we assessed the total number of screening measures each study used and how monetary bonuses were paid to participants (if at all). We note that this review was not meant to be systematic, but instead to provide a representative overview of common practices in online behavioral studies.

@@

Participants were eligible if they resided in the United States or Canada; participants from MTurk were recruited with the aid of CloudResearch services \cite{litman2017turkprime}. (Note: This study was conducted prior to the introduction of CloudResearch's new data quality filters \cite{cloudresearch_2020}). Following recent recommendations \cite{robinson2019tapped}, MTurk workers were not excluded based on work approval rate or number of previous jobs approved. No other exclusion criteria were applied during recruitment.

The data from multiple participants who completed the experiment were excluded prior to analysis. Three participants (all MTurk) were excluded due to missing data. In addition, we excluded 20 participants who disclosed that they had also completed the experiment on the other platform (MTurk: N=19, Prolific: N=1). This left a final sample of N=386 participants (MTurk: N=186, Prolific: N=200) for analysis. The demographics of the sample split by labour market is provided in Table S1. Notably, the participants recruited from MTurk were older ($M = 7.7 \ \text{yrs}, t = 6.567, p < 0.001$) and comprised of fewer women ($z = 6.567, p = 0.011$).

@@

The task was programmed in jsPsych \cite{deleeuw_2015_jspsych} and distributed using custom web-application software. The experiment code is available at \url{https://github.com/nivlab/ThreeArmJS}, and the web-software is available at \url{https://github.com/nivlab/nivturk}. A playable demo of the task is available at \url{https://nivlab.github.io/ThreeArmJS}.

@@

Prior to completing the reversal learning task, participants completed five self-report symptom measures. The symptom measures were selected for inclusion based on their frequency in clinical research, and for having an expected mixture of symmetric and asymmetric score distributions.

@@

@@

\subsection*{Bootstrapping Analysis}
\begin{figure}[H]
\includegraphics[width=16cm]{../figures/supp_01.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{The pattern of significant behavior-symptom correlations before (panel A) and after (panel B) screening compared to the resulting pattern when random subsets of participants (22\% of total, matched to screening using the infrequency measure) are removed (panel C). Panels A and B are reproduced from Figure \ref{fig:correlations} for convenience. Panel C: the fraction of significant correlations in 5000 bootstrapped samples. Panel D: The similarity of the pattern of correlations after removal of random subsets to that before and after screening using the infrequency measure. Similarity was calculated using the simple matching coefficient. Random removal subsets were significantly more similar to the ``No Screening" than to the ``Infrequency" screening datasets (t = 262.49, $p < 0.001$).}
\label{fig:random}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[width=16cm]{../figures/supp_02.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Posterior predictive checks for the risk-sensitive temporal difference learning model. \textit{Panel A:} Observed (black) and predicted (blue) learning curves averaged across the group. \textit{Panel B:} Observed versus predicted choice accuracy for each participant. \textit{Panel C:} Distribution of average predictive density across participants.}
\label{fig:posterior_check}
\end{figure}

@@

\subsection*{Participant Demographics}

\begin{table}[!h]
\centering
\small
\setlength{\tabcolsep}{15pt}
\begin{tabular}{ rrrrr }
\toprule
& \multicolumn{2}{c}{MTurk} & \multicolumn{2}{c}{Prolific} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
Total & \multicolumn{2}{c}{N=186} & \multicolumn{2}{c}{N=200} \\
\midrule
Age & N & \% & N & \% \\
\cmidrule(lr){1-5}
18-25   &      16 &   8.6 &       78 &  39.0 \\
26-35   &      76 &  40.9 &       69 &  34.5 \\
36-45   &      46 &  24.7 &       31 &  15.5 \\
46-55   &      22 &  11.8 &       13 &   6.5 \\
55+     &      26 &  14.0 &        9 &   4.5 \\
\midrule
Gender & N & \% & N & \% \\
\cmidrule(lr){1-5}
Female             &    83 &  44.6 &    112 &  56.0 \\
Male               &   103 &  55.4 &     85 &  42.5 \\
Other              &     0 &   0.0 &      3 &   1.5 \\
\midrule
Ethnicity & N & \% & N & \% \\
\cmidrule(lr){1-5}
Hispanic or Latino     &    15 &   8.1 &     10 &   5.0 \\
Not Hispanic or Latino &   168 &  90.3 &    183 &  91.5 \\
Rather not say         &     2 &   1.1 &      7 &   3.5 \\
Unknown                &     1 &   0.5 &      0 &   0.0 \\
\midrule
Race & N & \% & N & \% \\
\cmidrule(lr){1-5}
African American                 &    21 &  11.3 &      7 &   3.5 \\
Asian                                     &     5 &   2.7 &     53 &  26.5 \\
White                                     &   151 &  81.2 &    121 &  60.5 \\
Multiracial                               &     6 &   3.2 &      4 &   2.0 \\
Rather not say                            &     1 &   0.5 &     12 &   6.0 \\
\midrule
Use other platform & N & \% & N & \% \\
\cmidrule(lr){1-5}
Yes            &      71 &  38.2 &       28 &  14.0 \\
No             &     115 &  61.8 &      172 &  86.0 \\
\bottomrule
\end{tabular}
\captionsetup{width=0.88\textwidth}
\caption{The demographics of each sample by online labour market. On average, the samples were similar though the sample from Mechanical Turk was older (t = 6.567, p $<$ 0.001) and comprised of fewer women (z = 6.567, p = 0.011). Note: the demographics do not include 20 participants (MTurk: N=19; Prolific: N=1) excluded for participating in the study twice, once per platform.}
\label{tab:demographics}
\end{table}

@@

This result may be initially surprising, since one might expect choice hysteresis to result in more perseveration errors following contingency reversals, and since we observed in the previous section that C/IE responding was associated with more perseveration errors. However, choice hysteresis is adaptive in this probabilistic reversal-learning task. Because rewards in the task are probabilistic, once the reward-maximizing response option has been identified ignoring an occasional unrewarding outcome and instead performing the same response is optimal (until the next reversal occurs and is identified). Interestingly, participants engaging in C/IE responding were also numerically (though not significantly) less affected by previous outcomes, suggesting that their behavior was not more adaptive, but rather just more random.

>>>>

Here we wish to draw special attention to an underappreciated feature of psychiatric research using self-report symptom surveys. In such surveys, participants rate their endorsement of various psychiatric symptoms and, since most individuals in the general population tend to endorse no or few symptoms in many symptom domains, the resulting ground-truth symptom distributions tend to be heavily positively skewed \cite{lowe2008validation, tomitaka2018distributional}. In this situation, the assumption that C/IE responding merely increases unsystematic measurement noise becomes untenable. Because of the positive skew in the ground-truth symptom distribution, participants who respond carelessly to the symptom survey are more likely to report higher levels of symptom endorsement relative to participants who complete the survey attentively \cite{chandler2020participant, ophir2020turker, king2018random}. Consequently, unless C/IE survey responses are carefully identified and removed, a considerable proportion of putatively symptomatic individuals in an online sample may, in fact, be participants who have not engaged with the experiment with sufficient attention or effort.

@@

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low; that is, comparably higher scores are more likely for C/IE participants responding at random on a right skewed questionnaire. In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7; see Figure \ref{fig:distributions}). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct. Permutation testing confirmed that observed mean-shifts in symptom scores for C/IE participants were statistically significant for the majority of symptom measures (Table \ref{tab:validation}).

@@

Second, we quantified the degree to which participants responded to self-report symptom surveys in a stereotyped fashion; that is, we determined if participants exhibited patterns in their responses that were independent of the contents of the survey items. We fit a random-intercept item factor-analysis model \cite{maydeu2006random} to self-report data (see Methods), and for each participant we estimated an intercept parameter that quantified their bias towards using responses on the left or right side of the response scale, regardless of what that response signifies for a particular self-report measure (e.g., low on one symptom scale versus high on another). We observed a credible difference between the average value of this intercept for the two groups ($\Delta \text{intercept} = 0.67, 95\% \ \text{HDI} = [-0.78, -0.55]$), such that C/IE participants were biased towards using the right-half of survey response options. This translates to a tendency to endorse \emph{more severe} symptoms on the 7-up/7-down and GAD-7 scales (where the rightmost options indicate greater frequency of symptoms) but \emph{less extreme} symptoms or personality traits on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms or personality traits) despite these inventories measuring strongly correlated constructs (i.e. depression and anhedonia, anxiety and behavioral inhibition).

@@

\begin{figure}[!t]
\includegraphics[width=17cm]{../figures/main_03a.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Absolute Spearman rank correlations between task behavior (y-axis) and symptom measures (x-axis) under different regimes of data screening and participant exclusions. Only statistically significant correlations are shown ($p<0.05$ not corrected for multiple comparisons; all signed correlations are shown in the Supplement). Black Xs indicate significant correlations ablated under screening. No Screening = no exclusions; Accuracy Only = exclusions based on chance-level performance in the reversal-learning task; Infrequency Only = exclusions based on invalid or improbable responses to infrequency items; Both Types = exclusions based on the previous two measures. Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
\label{fig:correlations}
\end{figure}

@@

Next, we investigated how spurious correlations depended on sample size. To do so, we performed a bootstrapping analysis where we held fixed the proportion of participants engaging in C/IE responding (i.e. 5\%, 10\%, 15\%, 20\%) and increased the total number of participants. Across all analyses, we measured the correlation and between the 7-down depression scale and learning-rate asymmetry ($\kappa$), which we previously identified as likely exhibiting a spurious association. (The following results are not specific to learning-rate asymmetry and generalize to other pairs of variables; Figure \ref{fig:bootstrap_s1}).

The outputs of the bootstrapping analysis are presented in Figure \ref{fig:bootstrap}. We found that, although estimated correlation magnitudes were independent of sample size (x-axis, left panel), the absolute magnitude of the behavior-symptom correlation increased with the proportion of C/IE participants (different coloured circles, left panel). Crucially, we found false-positive rates for spurious correlations \emph{increased} with increases in sample size in our data for all but the smallest rates of C/IE responding (right panel). This runs counter to a common assumption that larger sample sizes are protective against spurious correlations because they serve to mitigate measurement error. Although this assumption is correct for unsystematic measurement error, it no longer holds in the regime of systematic measurement error (where larger sample sizes reduce the variance of estimates, but do not alter their bias). Instead, our results suggest that, except for low rates of C/IE responding, the false-positive rate for behavioral-symptom correlations will become increasingly inflated as the sample size increases.

@@

Notwithstanding the above, one response to this legitimate concern is to take a graded approach to screening and excluding participants \cite{Kim2018-ev}. That is, participants could be screened with respect to a multitude of measures and only the consistently flagged participants be removed, thereby reducing the risk of inducing bias. Another possibility is to use sensitivity analysis as an alternative to exclusion, testing whether full-sample observed correlations are robust to the exclusion of participants flagged by measures of C/IE responding. We note that the strict screening approach used in the present study did not preclude us from identifying symptomatic participants or behavior-symptom correlations. Indeed, we found in our sample roughly 10\% of participants endorsing symptoms consistent with clinical levels of depression, and approximately 20\% consistent with clinical levels of acute anxiety. These estimates are within the realm of epidemiological norms \cite{kessler2012twelve, lowe2008validation, yarrington2021impact}. (We should note, however, that some studies have found elevated rates of psychiatric symptomology in online participants even after controlling for C/IE responding \cite{ophir2020turker}.) We also observed some positive correlations between anxiety and choice behavior that were consistent with effects found in previous literature \cite{huang2017computational, harle2017anhedonia, garrett2018updating}. For example, we found higher lose-shift rates and higher learning rates following negative prediction errors correlated with self-reported anxiety. This suggests that the screening methods we employed were not so aggressive as to attenuate behavior-symptom correlations that would be expected from the literature.

@@

    \item When collecting self-report questionnaire data, include attention-check items that flag participants who may be engaging in C/IE responding. We recommend following best-practice guidelines in using infrequency-item checks rather than instructed-item checks, as multiple studies have now shown that online participants are habituated to and circumvent the latter  \cite{barends2019noncompliant, thomas2017validity, hauser2016attentive}. Participants flagged by suspicious responses on attention-check items should either be excluded from further analysis, or assessed using sensitivity analyses to ensure that observed full-sample correlations are robust to their exclusion.

@@

To characterize common data screening practices in online computational psychiatry studies, we performed a narrative literature review \cite{grant2009typology}. We identified studies for inclusion through searches on Google Scholar using permutations of query terms related to online labour platforms (e.g. ``mechanical turk", ``prolific", ``online"), experimental paradigms (e.g. ``experiment", ``cognitive control", ``reinforcement learning"), and symptom measures (e.g. ``psychiatry", ``mental illness", ``depression"). We note that it was not feasible to conduct a systematic review, which would require using a publication database other than Google Scholar (i.e. this is because search results from Google Scholar are tailored to individual users and are therefore by definition unsystematic and unreproducible). Our science librarian nevertheless recommended we use Google Scholar because it performs full-text search, which was necessary here to identify papers; that is, recruitment methods (e.g. ``Mechanical Turk'') are infrequently reported in the title or abstract of a paper. We included studies that (a) recruited participants online through a labour platform, (b) measured behavior on at least one experimental task, and (c) measured responses on at least one self-report symptom measure. Through this approach, we identified for inclusion 49 studies spanning 2015 through 2020. The complete list of studies, and search terms used to find them, are included in the Github repository for this study.

Two of the authors (S.Z., D.B.) then evaluated whether and how each of these studies performed data quality screening for both the collected task and self-report data. Specifically, we confirmed whether a study had performed a particular type of data screening. The categories were determined based on previous taxonomies of screening methods (e.g. \cite{curran2016methods}). In addition, we assessed the total number of screening measures each study used and how monetary bonuses were paid to participants (if at all). We note that this review was not meant to be systematic, but instead to provide a representative overview of common practices in online behavioral studies.

@@

Participants were eligible if they resided in the United States or Canada; participants from MTurk were recruited with the aid of CloudResearch services \cite{litman2017turkprime}. (Note: This study was conducted prior to the introduction of CloudResearch's newest data quality filters \cite{cloudresearch_2020}). Following recent recommendations \cite{robinson2019tapped}, MTurk workers were not excluded based on work approval rate or number of previous jobs approved. No other exclusion criteria were applied during recruitment. It is important to note that both CloudResearch and Prolific internally use a number of tools (e.g. IP-address screening) to filter out the lowest quality participants. In addition, our custom experiment delivery software (NivTurk; see below) has bot-checking functionality built into it. We are confident that our study is not strongly affected by participants using software to automatically complete the experiment.

The data from multiple participants who completed the experiment were excluded prior to analysis. Three participants (all MTurk) were excluded due to missing data. In addition, we excluded 20 participants who disclosed that they had also completed the experiment on the other platform. This left a final sample of N=386 participants (MTurk: N=186, Prolific: N=200) for analysis. The demographics of the sample split by labour market is provided in Table S1. Notably, the participants recruited from MTurk were older ($M = 7.7 \ \text{yrs}, t = 6.567, p < 0.001$) and comprised of fewer women ($z = 6.567, p = 0.011$).

@@

The task was programmed in jsPsych \cite{deleeuw_2015_jspsych} and distributed using custom web-application software. The experiment code is available at \url{https://github.com/nivlab/sciops}, and the web-software is available at \url{https://github.com/nivlab/nivturk}. A playable demo of the task is available at \url{https://nivlab.github.io/jspsych-demos/tasks/3arm/experiment.html}.

@@

Prior to completing the reversal learning task, participants completed five self-report symptom and personality-trait measures. The symptom measures were selected for inclusion based on their frequency in clinical research, and for having an expected mixture of symmetric and asymmetric score distributions.

@@

\subparagraph{Task-based screening variables}

@@

\subsection*{Signed correlation analysis}

\begin{figure}[H]
\includegraphics[width=17cm]{../figures/supp_01.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Signed Spearman rank correlations between task behavior (y-axis) and symptom measures (x-axis) under different regimes of data screening and participant exclusions. Only statistically significant correlations are shown ($p<0.05$ not corrected for multiple comparisons). Black Xs indicate significant correlations ablated under screening. No Screening = no exclusions; Accuracy Only = exclusions based on chance-level performance in the reversal-learning task; Infrequency Only = exclusions based on invalid or improbable responses to infrequency items; Both Types = exclusions based on the previous two measures. Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
\label{fig:correlations_s1}
\end{figure}

\subsection*{Bootstrapping Analysis}
\begin{figure}[H]
\includegraphics[width=16cm]{../figures/supp_02.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{The pattern of significant behavior-symptom correlations before (panel A) and after (panel B) screening compared to the resulting pattern when random subsets of participants (22\% of total, matched to screening using the infrequency measure) are removed (panel C). Panels A and B are reproduced from Figure \ref{fig:correlations} for convenience. Panel C: the fraction of significant correlations in 5000 bootstrapped samples. Panel D: The similarity of the pattern of correlations after removal of random subsets to that before and after screening using the infrequency measure. Similarity was calculated using the simple matching coefficient. Random removal subsets were significantly more similar to the ``No Screening" than to the ``Infrequency" screening datasets (t = 262.49, $p < 0.001$).}
\label{fig:random}
\end{figure}

\subsection*{The relationship between sample size and false positive rates generalize to other sets of variables}

\begin{figure}[H]
\includegraphics[width=16cm]{../figures/supp_03.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{False positive rates for spurious correlations \emph{increase} with sample size. \textit{Left:} Bootstrapped Spearman correlations between inverse temperature ($\beta$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. The grey error bars indicate 95\% bootstrap confidence intervals. The thick dashed lines indicate the threshold for statistical significance for the Spearman correlation at the corresponding sample size. Markers are jittered along the x-axis for legibility. \textit{Right:} False positive rates for inverse temperature ($\beta$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. False positive rate was calculated as the proportion of bootstrap samples in which the correlation between $\beta$ and 7-down was statistically significant. The horizontal dotted line denotes the expected false positive rate at $\alpha = .05$.}
\label{fig:bootstrap_s1}
\end{figure}

\subsection*{The relationship between sample size and true positive rates for true correlations}

\begin{figure}[H]
\includegraphics[width=16cm]{../figures/supp_04.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{True correlations are independent of the proportion of C/IE participants in the sample. \textit{Left:} Bootstrapped Spearman correlations between learning rate asymmetry ($\kappa$) and anxiety scores (GAD-7) as a function of sample size and proportion of C/IE participants. The grey error bars indicate 95\% bootstrap confidence intervals. The thick dashed lines indicate the threshold for statistical significance for the Spearman correlation at the corresponding sample size. Markers are jittered along the x-axis for legibility. \textit{Right:} True positive rates for learning rate asymmetry ($\kappa$) and anxiety scores (GAD-7) as a function of sample size and proportion of C/IE participants. True positive rate was calculated as the proportion of bootstrap samples in which the correlation between $\kappa$ and GAD-7 was statistically significant. The horizontal dotted line denotes the expected false positive rate at $\alpha = .05$.}
\label{fig:bootstrap_s2}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[width=16cm]{../figures/supp_05.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Posterior predictive checks for the risk-sensitive temporal difference learning model. \textit{Panel A:} Observed (black) and predicted (blue) learning curves averaged across the group. \textit{Panel B:} Observed versus predicted choice accuracy for each participant. \textit{Panel C:} Distribution of average predictive density across participants.}
\label{fig:posterior_check}
\end{figure}

@@

\subsection*{Participant demographics}

\begin{table}[!h]
\centering
\small
\setlength{\tabcolsep}{15pt}
\begin{tabular}{ rrrrr }
\toprule
& \multicolumn{2}{c}{MTurk} & \multicolumn{2}{c}{Prolific} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
Total & \multicolumn{2}{c}{N=186} & \multicolumn{2}{c}{N=200} \\
\midrule
Age & N & \% & N & \% \\
\cmidrule(lr){1-5}
18-25   &      16 &   8.6 &       78 &  39.0 \\
26-35   &      76 &  40.9 &       69 &  34.5 \\
36-45   &      46 &  24.7 &       31 &  15.5 \\
46-55   &      22 &  11.8 &       13 &   6.5 \\
55+     &      26 &  14.0 &        9 &   4.5 \\
\midrule
Gender & N & \% & N & \% \\
\cmidrule(lr){1-5}
Female             &    83 &  44.6 &    112 &  56.0 \\
Male               &   103 &  55.4 &     85 &  42.5 \\
Other              &     0 &   0.0 &      3 &   1.5 \\
\midrule
Ethnicity & N & \% & N & \% \\
\cmidrule(lr){1-5}
Hispanic or Latino     &    15 &   8.1 &     10 &   5.0 \\
Not Hispanic or Latino &   168 &  90.3 &    183 &  91.5 \\
Rather not say         &     2 &   1.1 &      7 &   3.5 \\
Unknown                &     1 &   0.5 &      0 &   0.0 \\
\midrule
Race & N & \% & N & \% \\
\cmidrule(lr){1-5}
African American                 &    21 &  11.3 &      7 &   3.5 \\
Asian                                     &     5 &   2.7 &     53 &  26.5 \\
White                                     &   151 &  81.2 &    121 &  60.5 \\
Multiracial                               &     6 &   3.2 &      4 &   2.0 \\
Rather not say                            &     1 &   0.5 &     12 &   6.0 \\
\midrule
Use other platform & N & \% & N & \% \\
\cmidrule(lr){1-5}
Yes            &      71 &  38.2 &       28 &  14.0 \\
No             &     115 &  61.8 &      172 &  86.0 \\
\bottomrule
\end{tabular}
\captionsetup{width=0.88\textwidth}
\caption{The demographics of each sample by online labour market. On average, the samples were similar though the sample from Mechanical Turk was older (t = 6.567, p $<$ 0.001) and comprised of fewer women (z = 6.567, p = 0.011). Note: the demographics do not include 20 participants excluded for participating in the study twice, once per platform.}
\label{tab:demographics}
\end{table}

\subsection*{Careless participants show different behaviors on the reversal learning task}

\begin{table}[H]
\centering
\begin{tabular}{lrrr}
\toprule
{} &  Attentive &    C/IE &   $t$-value \\
\midrule
Acc      &      0.587 &   0.532 &  $4.008^*$  \\
Pts      &     50.163 &  47.729 &  $2.376^*$  \\
WS       &      0.898 &   0.776 &  $5.387^*$  \\
LS       &      0.609 &   0.751 &  $-5.335^*$ \\
Pers     &      0.245 &   0.259 &  $-1.505$ \ \\
$\beta$  &      6.754 &   4.082 &  $5.404^*$  \\
$\eta_p$ &      0.643 &   0.551 &  $2.846^*$  \\
$\eta_n$ &      0.738 &   0.784 & $-1.516$ \  \\
$\kappa$ &     -0.069 &  -0.218 &  $3.729^*$  \\
\bottomrule
\end{tabular}
\captionsetup{width=0.88\textwidth}
\caption{Measures of task behavior compared between attentive and C/IE participants. Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry. * Denotes statistical significance ($\alpha=0.05$, not corrected for multiple comparisons).}
\end{table}

@@

This result may be initially surprising, since one might expect choice hysteresis to result in more perseveration errors following contingency reversals. However, choice hysteresis is adaptive in this probabilistic reversal-learning task. Because rewards in the task are probabilistic, once the reward-maximizing response option has been identified ignoring an occasional unrewarding outcome and instead performing the same response is optimal (until the next reversal occurs and is identified). Interestingly, participants engaging in C/IE responding were also numerically (though not significantly) less affected by previous outcomes, suggesting that their behavior was not more adaptive, but rather just more random.
